---
title: 'STA305H5S: Project'
author: "Bruce Liu, Chunxu Hou, jinke lyu, Yiran Zhai"
date: "Winter 2026"
output:
  pdf_document:
    number_sections: true
    latex_engine: xelatex
  html_document:
    df_print: paged
  word_document: default
spacing: double
header-includes:
- \usepackage{fancyhdr}
- \pagestyle{fancy}
- "\\fancyhead[L]{STA305H5-Experimental Design}"
- \fancyfoot[CO,CE]{}
- \fancyfoot[LE,RO]{\thepage}
- \usepackage{setspace}\doublespacing
---


```{=tex}
\newpage 
\tableofcontents 
\newpage
```

```{r setup, include=FALSE}
library(scidesignR)
data(fertdat)

model <- lm(fert ~ trt, data = fertdat)
res <- residuals(model)
```

# Part 1 
This study investigates whether two different fertilizers lead to different levels of wheat yield. Understanding such differences is important in agricultural practice, as selecting a more effective fertilizer can improve productivity and resource efficiency.
The objective is to determine whether any observed difference in average yield between Fertilizer A and Fertilizer B is due to a real treatment effect or simply random variation. This will be assessed using statistical methods, including exploratory analysis and hypothesis testing.

## Research Question
Determine whether there is a statistically significant difference in mean wheat yield between Fertilizer A and Fertilizer B.

## Experimental Units
The experimental units are individual plots (or plants), each receiving one type of fertilizer treatment. These units are assumed to be independent.

## Response Variable
The response variable is wheat yield, measured for each experimental unit. This variable reflects the effectiveness of the fertilizer treatment.

## Treatment Factor
The treatment factor is fertilizer type, with two levels: Fertilizer A and Fertilizer B. The goal is to compare these two treatments to assess their impact on wheat yield.

## Exploratory Data Analysis
The sample mean wheat yield for Fertilizer A is 22.18, while for Fertilizer B is 20.48, suggesting a slightly higher average yield under Fertilizer A. However, summary statistics alone do not indicate whether this difference is meaningful.
The boxplot below compares the distributions of wheat yield for the two fertilizer groups. It shows that both groups have similar medians and comparable variability, with no clear separation between them.

```{r, echo=FALSE, out.width="60%", fig.align="center"}
boxplot(fert ~ trt, data = fertdat,
        xlab = "Fertilizer Type",
        ylab = "Wheat Yield",
        main = "Comparison of Wheat Yield by Fertilizer Type")
```

## Assumption Checks
To validate the use of a two-sample t-test, we examine the assumptions of normality and equal variance.
The normality assumption is assessed using a Q-Q plot of the residuals and the Shapiro–Wilk test. The plot shows points approximately along the reference line, and the test yields a p-value of 0.8418, indicating no evidence against normality.

```{r, echo=FALSE, out.width="60%", fig.align="center"}
qqnorm(res)
qqline(res)
```{r, echo=FALSE, fig.cap="Normal Q-Q plot of residuals"}
```
The constant variance assumption is assessed using a residuals versus fitted values plot. The residuals appear randomly scattered around zero with no clear pattern, suggesting that the variance is approximately constant.

```{r, echo=FALSE, out.width="60%", fig.align="center"}
plot(fitted(model), res,
     xlab = "Fitted Values",
     ylab = "Residuals",
     main = "Residuals vs Fitted Plot")
abline(h = 0, col = "red")
```{r, echo=FALSE, fig.cap="Residuals vs fitted values"}
```
In addition, an F-test for equality of variances gives a p-value of 0.217, which supports the assumption that the two groups have equal variances.


## Hypothesis Test
To formally assess whether the two fertilizers lead to different average wheat yields, a two-sample t-test is conducted.
Let $\mu_A$ and $\mu_B$ denote the mean wheat yield under Fertilizer A and Fertilizer B, respectively. The hypotheses are:
Null hypothesis: $\mu_A$ = $\mu_B$  
Alternative hypothesis: $\mu_A$ $\neq$ $\mu_B$  
The sample results indicate that the mean yield for Fertilizer A is slightly higher than that for Fertilizer B, with an estimated difference (A − B) of 1.70.
A 95% confidence interval for the difference in means is (-5.57, 8.97). Since this interval includes zero, it suggests that the true difference may be zero.
The p-value for the test is 0.6137, which is substantially larger than the significance level of 0.05.

## Results and Interpretation
Since the p-value (0.6137) is greater than 0.05, we fail to reject the null hypothesis. This indicates that there is no statistically significant evidence of a difference in mean wheat yield between the two fertilizers.
The confidence interval (-5.57, 8.97) also contains zero, reinforcing this conclusion.
Although the observed difference in means (1.70) suggests that Fertilizer A may have a slightly higher yield, this difference is small relative to the variability in the data.
Overall, there is insufficient evidence to conclude that Fertilizer A and Fertilizer B produce different mean wheat yields.

# Part 2

## Effect Size

The effect size in this study is about 0.30. This indicates that the average wheat yield under Fertilizer A is higher than that under Fertilizer B in the observed sample as shown above. However, an effect size of 0.30 also suggests that the treatment difference is small relative to the variability in the data.

```{r, echo = FALSE}
# Effect Size

library(scidesignR)
data("fertdat")

fert_A <- fertdat$fert[fertdat$trt == "A"]
fert_B <- fertdat$fert[fertdat$trt == "B"]

mean_A <- mean(fert_A)
mean_B <- mean(fert_B)
delta <- mean_A - mean_B

sd_A <- sd(fert_A)
sd_B <- sd(fert_B)

sp <- sqrt(((length(fert_A) - 1) * sd_A^2 + (length(fert_B) - 1) * sd_B^2) / 
             (length(fert_A) + length(fert_B) - 2))

ES <- delta / sp
```

## Sample size
To design the study, we calculated the required sample size for a two-sample t-test with 80% power at the 5% significance level with the exact method "power.t.test()". The required sample size within each group is: 
```{r, echo = FALSE}
# Exact Method
power <- 0.8
alpha <- 0.05

exact_n <- power.t.test(
  power = power,
  delta = delta,
  sd = sp, 
  sig.level = alpha
)
exact_n
# Round up the sample size
req_n <- ceiling(exact_n$n)
req_n
```

This output gives the required sample size of 174.4104 in each group. Since the sample size must be a whole number, I rounded up the size to 175. Therefore, this study would require 175 experimental units in each fertilizer group in order to achieve about 80% power.

In order to check the result, we also calculated the power by simulation. We repeatedly generated two sample size of 175 in each group, one from the normal distribution with mean equal to the observed mean of Fertilizer A and one from a normal distribution with mean equal to the observed mean of Fertilizer B.
```{r, echo = FALSE}
# Simulation Power check 
set.seed(1)
alpha <- 0.05

pvals <- replicate(10000,
                   t.test(
                     rnorm(req_n, mean = mean_A, sd = sd_A),
                     rnorm(req_n, mean = mean_B, sd = sd_B),
                     var.equal = TRUE)$p.value)
sim_p <- sum((pvals <= alpha) / 10000)
sim_p
```
The simulation estimates a power of 0.8083, which is very close to the target power of 0.8. It proves that the required sample size of 175 using the exact method is correct. 

## Sensetivity to Effect Size and Varibility

In order to examine how the required sample size would change if effect size is smaller, or variability is larger. 

Firstly, when the treatment difference between fertilizer A and fertilizer B is smaller and standard deviation remains the same. The new required sample size within each group will be:

```{r, echo = FALSE}
# When the effect side is smaller
delta_small <- 1
power <- 0.8
alpha <- 0.05

exact_n_smallD <- power.t.test(
  power = power,
  delta = delta_small,
  sd = sp, 
  sig.level = alpha
)
exact_n_smallD
# Round up the sample size
req_n_smallD <- ceiling(exact_n_smallD$n)
req_n_smallD
```

Now considering a larger standard deviation while the difference between two treatment remains the same. The new required sample size is: 

```{r, echo = FALSE}
# When the variability is larger
SD_Large <- 10
power <- 0.8
alpha <- 0.05

exact_n_largeSD <- power.t.test(
  power = power,
  delta = delta_small,
  sd = SD_Large, 
  sig.level = alpha
)
exact_n_largeSD
# Round up the sample size
req_n_largeSD <- ceiling(exact_n_largeSD$n)
req_n_largeSD
```

Therefore, when the effect size is reduced from 1.7 to 1, he required sample size increased from 175 to 503. When the standard deviation increased from 5.65 to 10, the required sample size increased from 175 to 1571. Both change result a increase in the required sample size. 



# Part 3

## 3A: One-way ANOVA

### Description

```{r}
library(scidesignR)
data(painstudy)
summary(painstudy)
```

The dataset *painstudy* contains observations of pain levels under different treatment groups. The objective is to determine whether the mean pain levels differ across treatments.


### Check Assumptions

```{r}
aov.pain <- aov(pain ~ trt, data = painstudy)
summary(aov.pain)
```

```{r}
# Checking Assumption: Constant Variance
plot(aov.pain$fitted.values, aov.pain$residuals,
     ylab = "Residuals", xlab = "Fitted",
     main = "One-way ANOVA")
abline(h = 0)
```

```{r}
# Checking Assumption: Normality
qqnorm(aov.pain$residuals, main = "Normal Q-Q Plot")
qqline(aov.pain$residuals)
```

To evaluate the assumptions of the ANOVA model, diagnostic plots were examined. The residuals appear randomly scattered, suggesting constant variance, and the Q-Q plot indicates approximate normality.


### Interpret Results

The null hypothesis is that all treatment means are equal:

H₀: μ₁ = μ₂ = μ₃  

The alternative hypothesis is that at least one treatment mean differs:

Hₐ: Not all μ are equal.

Based on the ANOVA output, the p-value is less than 0.001.

Since the p-value is smaller than the significance level (α = 0.05), we reject the null hypothesis.

```{r}
TukeyHSD(aov(pain ~ trt, data = painstudy))
```

Further analysis using Tukey’s post-hoc test shows that all pairwise differences between treatments are statistically significant.

---

### Conclusion

There is sufficient evidence to conclude that treatment has a statistically significant effect on pain levels. Among the treatments, Treatment B has the highest mean response, followed by Treatment A, while Treatment C has the lowest.

## 3B: Randomized Block Design

### Description

```{r}
data(shoedat_obs)
summary(shoedat_obs)
```

```{r}
shoe_long <- data.frame(
  boy = rep(shoedat_obs$boy, 2),
  treatment = rep(c("A", "B"), each = nrow(shoedat_obs)),
  side = c(shoedat_obs$sideA, shoedat_obs$sideB),
  wear = c(shoedat_obs$wearA, shoedat_obs$wearB)
)

shoe_long$boy <- factor(shoe_long$boy)
shoe_long$treatment <- factor(shoe_long$treatment)
```

### Check Assumptions

```{r}
rbd_model <- aov(wear ~ treatment + boy, data = shoe_long)
summary(rbd_model)
```

```{r}
plot(rbd_model$fitted.values, rbd_model$residuals,
     xlab = "Fitted values",
     ylab = "Residuals",
     main = "Residuals vs Fitted")
abline(h = 0)
```

```{r}
qqnorm(rbd_model$residuals, main = "Normal Q-Q Plot")
qqline(rbd_model$residuals)
```

### Interpret Results

```{r}
unblocked_model <- aov(wear ~ treatment, data = shoe_long)
summary(unblocked_model)
```

The null hypothesis is:

H₀: The mean wear is the same for both treatments.

The alternative hypothesis is:

Hₐ: The mean wear differs between the two treatments.

Based on the ANOVA output, the p-value for treatment is less than 0.05, so we reject H₀.

Compared with the unblocked analysis, including boy as a blocking factor reduces the residual variation, which means part of the variability is explained by differences between boys.

### Conclusion

There is sufficient evidence to conclude that treatment has a significant effect on wear. Using boy as a blocking factor improves the precision of the treatment comparison by accounting for individual differences.

## 3C: Latin Square

### Description

```{r}
data(BR_LatSq)
summary(BR_LatSq)
```

In the dataset BR_LatSq, the response variable is Value, the treatment variable is Treat, and the two blocking factors are Dogs and Weeks.

This is a Latin square design because it controls for two possible sources of nuisance variation at the same time. In this experiment, Dogs and Weeks are treated as blocking factors, while the main objective is to evaluate whether Treat has a significant effect on the response.

### Check Assumptions

```{r}
BR_LatSq$Dogs <- as.factor(BR_LatSq$Dogs)
BR_LatSq$Weeks <- as.factor(BR_LatSq$Weeks)
BR_LatSq$Treat <- as.factor(BR_LatSq$Treat)
```

```{r}
lat.model <- lm(Value ~ Treat + as.factor(Dogs) + as.factor(Weeks), data = BR_LatSq)
anova(lat.model)
```

```{r}
qqnorm(lat.model$residuals, main = "Q-Q Plot for BR_LatSq")
qqline(lat.model$residuals)
```

```{r}
plot(lat.model$fitted.values, lat.model$residuals,
     ylab = "Residuals", xlab = "Fitted",
     main = "Residual Plot")
abline(h = 0)
```

The Q-Q plot suggests that the residuals are approximately normally distributed. The residual plot does not show any strong pattern, so the assumption of constant variance appears reasonable. Overall, there is no strong evidence that the model assumptions are violated.

### Interpret Results

The null hypothesis is:

H₀: All treatment means are equal.

The alternative hypothesis is:

Hₐ: At least one treatment mean differs.

Based on the ANOVA table, the p-value for Treat is approximately 0.1362.

Since this p-value is greater than the significance level of 0.05, we fail to reject the null hypothesis.

In addition, the blocking factors Dogs and Weeks are also not statistically significant, with p-values of approximately 0.2198 and 0.8878, respectively.

### Conclusion

There is insufficient evidence to conclude that the treatment has a significant effect on the response variable after accounting for the blocking factors Dogs and Weeks.

This suggests that, in this experiment, neither the treatment nor the two blocking factors explain a statistically significant amount of variation in the response.


The data from the R package \textit{coronavirus} has been reconstructed to have the first column representing recordings 
over time from January to June 2020, and the next three columns representing the total number of confirmed cases, deaths 
and recoveries in Canada. B


```{r cars}
summary(cars)
```
 
 
Based on this data, it is possible to estimate the number of individuals infected daily by,
 
```{r, echo=FALSE}
plot(pressure)
```

A simple linear regression is given by
$$
y=\beta_0 + \beta_1 x + \varepsilon
$$


# Part 4

```{r}
data(wtlossdat)
summary(wtlossdat)
```


There are three factors in this study, they are A, B, and C. 

A is the level of food diary with two level 1 and -1. 
B is the level of increasing activity with two level 1 and -1. 
C is the level of home visit with two level 1 and -1. 

The response variable is y which means weight loss(kg).


```{r}
# estimate main effects and interactions method 1
complete_model <- lm(y ~ A*B*C, data = wtlossdat)

coef_estimates <- coef(complete_model)

effect_A   <- 2 * coef_estimates["A"]
effect_B   <- 2 * coef_estimates["B"]
effect_C   <- 2 * coef_estimates["C"]

effect_AB  <- 2 * coef_estimates["A:B"]
effect_AC  <- 2 * coef_estimates["A:C"]
effect_BC  <- 2 * coef_estimates["B:C"]

effect_ABC <- 2 * coef_estimates["A:B:C"]

effects_summary <- data.frame(
  Effect = c("A", "B", "C", "AB", "AC", "BC", "ABC"),
  Estimate = c(effect_A, effect_B, effect_C,
               effect_AB, effect_AC, effect_BC, effect_ABC)
)

effects_summary
```

```{r}
# estimate main effects and interactions method 2
#library(FrF2)

wt.model <- lm(y ~ A*B*C, data = wtlossdat)
cubePlot(wt.model, "A", "B", "C", main = "Cube plot for wtlossdat")
mean_A_pos <- (1.0 + 2.6 - 0.4 + 0.4) / 4
mean_A_neg  <- (1.1 + 1.8 - 0.3 - 1.1) / 4
effect_A_mean <- mean_A_pos - mean_A_neg

mean_B_pos <- (-0.3 - 1.1 - 0.4 + 0.4) / 4
mean_B_neg  <- (1.1 + 1.8 + 1.0 + 2.6) / 4
effect_B_mean <- mean_B_pos - mean_B_neg

mean_C_pos <- (1.8 - 1.1 + 2.6 + 0.4) / 4
mean_C_neg <- (1.1 - 0.3 + 1.0 - 0.4) / 4
effect_C_mean <- mean_C_pos - mean_C_neg

mean_same_symbol_AB  <- (1.1 + 1.8 - 0.4 + 0.4) / 4
mean_different_symbol_AB <- (-0.3 - 1.1 + 1.0 + 2.6) / 4
effect_AB_mean <- mean_same_symbol_AB - mean_different_symbol_AB

mean_same_symbol_AC  <- (1.1 - 0.3 + 2.6 + 0.4) / 4
mean_different_symbol_AC <- (1.8 - 1.1 + 1.0 - 0.4) / 4
effect_AC_mean <- mean_same_symbol_AC - mean_different_symbol_AC

mean_same_symbol_BC  <- (1.1 - 1.1 + 1.0 + 0.4) / 4
mean_different_symbol_BC <- (1.8 - 0.3 + 2.6 - 0.4) / 4
effect_BC_mean <- mean_same_symbol_BC - mean_different_symbol_BC

mean_product_ABC_pos  <- (1.8 - 0.3 + 1.0 + 0.4) / 4
mean_product_ABC_neg <- (1.1 - 1.1 + 2.6 - 0.4) / 4
effect_ABC_mean <- mean_product_ABC_pos - mean_product_ABC_neg

effects.mean <- data.frame(
  Effect = c("A","B","C","AB","AC","BC","ABC"),
  Estimate = c(effect_A_mean, effect_B_mean, effect_C_mean,
               effect_AB_mean, effect_AC_mean, effect_BC_mean, effect_ABC_mean)
)
effects.mean
```

# main effects/interaction plots to support interpretation

# A mean

mean_A_pos <- (1.0 + 2.6 - 0.4 + 0.4) / 4
mean_A_neg  <- (1.1 + 1.8 - 0.3 - 1.1) / 4

# B mean

mean_B_pos <- (-0.3 - 1.1 - 0.4 + 0.4) / 4
mean_B_neg  <- (1.1 + 1.8 + 1.0 + 2.6) / 4

# C mean

mean_C_pos <- (1.8 - 1.1 + 2.6 + 0.4) / 4
mean_C_neg <- (1.1 - 0.3 + 1.0 - 0.4) / 4

par(mfrow = c(1, 3))

# plot A

plot(c(-1, 1), c(mean_A_neg, mean_A_pos), type = "b", lwd = 3, pch = 19, col = "blue",
     main = "effect_A_mean", xlab = "A", ylab = "Weight Loss")

# plot B

plot(c(-1, 1), c(mean_B_neg, mean_B_pos), type = "b", lwd = 3, pch = 19, col = "red",
     main = "effect_B_mean", xlab = "B", ylab = "Weight Loss")

# plot C

plot(c(-1, 1), c(mean_C_neg, mean_C_pos), type = "b", lwd = 3, pch = 19, col = "green",
     main = "effect_C_mean", xlab = "C", ylab = "Weight Loss")

par(mfrow = c(1, 1))

# AB interaction mean
mean_same_symbol_AB <- (1.1 + 1.8 - 0.4 + 0.4)/4
mean_different_symbol_AB <- (-0.3 -1.1 +1.0 +2.6)/4

A_neg_B_neg <- (1.1 + 1.8)/2
A_neg_B_pos <- (-0.3 -1.1)/2
A_pos_B_neg <- (1.0 + 2.6)/2
A_pos_B_pos <- (-0.4 + 0.4)/2

# AC interaction mean
mean_same_symbol_AC <- (1.1 -0.3 +2.6 +0.4)/4
mean_different_symbol_AC <- (1.8 -1.1 +1.0 -0.4)/4

A_neg_C_neg <- (1.1 -0.3)/2
A_neg_C_pos <- (1.8 -1.1)/2
A_pos_C_neg <- (1.0 -0.4)/2
A_pos_C_pos <- (2.6 + 0.4)/2

# BC interaction mean
mean_same_symbol_BC <- (1.1 -1.1 +1.0 +0.4)/4
mean_different_symbol_BC <- (1.8 -0.3 +2.6 -0.4)/4

B_neg_C_neg <- (1.1 +1.0)/2
B_neg_C_pos <- (1.8 +2.6)/2
B_pos_C_neg <- (-0.3 -0.4)/2
B_pos_C_pos <- (-1.1 +0.4)/2

par(mfrow=c(1,3))

# AB interaction plot
plot(c(-1,1), c(A_neg_B_neg, A_pos_B_neg), type="b", lwd=3, pch=19, col="red", main="AB Interaction", xlab="A", ylab="y")
lines(c(-1,1), c(A_neg_B_pos, A_pos_B_pos), type="b", lwd=3, pch=19, col="blue")

# AC interaction plot
plot(c(-1,1), c(A_neg_C_neg, A_pos_C_neg), type="b", lwd=3, pch=19, col="green",main="AC
Interaction", xlab="A", ylab="y")
lines(c(-1,1), c(A_neg_C_pos, A_pos_C_pos), type="b", lwd=3, pch=19, col="purple")

# BC interaction plot
plot(c(-1,1), c(B_neg_C_neg, B_pos_C_neg), type="b", lwd=3, pch=19, col="orange", main="BC Interaction", xlab="B", ylab="y")
lines(c(-1,1), c(B_neg_C_pos, B_pos_C_pos), type="b", lwd=3, pch=19, col="brown")

par(mfrow=c(1,1))

# ABC interaction mean
A_neg_B_neg_C_neg = 1.1
A_neg_B_pos_C_neg = -0.3
A_pos_B_neg_C_neg = 1.0
A_pos_B_pos_C_neg = -0.4
A_neg_B_neg_C_pos = 1.8
A_pos_B_neg_C_pos = 2.6
A_neg_B_pos_C_pos = -1.1
A_pos_B_pos_C_pos = 0.4

par(mfrow = c(2,1))

# ABC interaction plot
# where C = 1
plot(c(-1, 1), c(A_neg_B_neg_C_pos, A_pos_B_neg_C_pos), type = "b", lwd = 3, pch = 19, col = "red",main = "ABC Interaction (C = 1)", xlab = "A", ylab = "Weight Loss", ylim = c(-1.5, 3))
lines(c(-1, 1), c(A_neg_B_pos_C_pos, A_pos_B_pos_C_pos), type = "b", lwd = 3, pch = 19, col = "blue")

# where C = -1
plot(c(-1, 1), c(A_neg_B_neg_C_neg, A_pos_B_neg_C_neg), type = "b", lwd = 3, pch = 19, col = "red",main = "ABC Interaction (C = -1)", xlab = "A", ylab = "Weight Loss", ylim = c(-1.5, 3))
lines(c(-1, 1), c(A_neg_B_pos_C_neg, A_pos_B_pos_C_neg), type = "b", lwd = 3, pch = 19, col = "blue")

par(mfrow = c(1,1))

# recommendation

# effect A (Food diary)

The main effect of A is positive. Keeping a food diary (which means 1) increases average weight loss compared to not keeping a diary (which means -1).

So, factor A is beneficial for weight reduction.

# effect B (Increasing activity)

The main effect of B is negative. Increasing activity (which means 1) decreases average weight loss in this experiment.

So, factor B is disadvantageous for weight reduction.

# effect C (Home visit)

The main effect of C is positive. Receiving home visit support (which means 1) increases average weight loss compared to no home visit (which means -1).

So,factor C is beneficial for weight reduction.

# AB Interaction

According to the plot, the interaction of AB is small which inmplies the effect of food diary (A) barely depends on whether increasing activity (B) is increased or not. 

So, there is not exist a strong relationship between A and B

# AC Interaction

According to the plot, the interaction of AC is weak positive which implies the benefit of a food diary(A) slightly increases when home visit(c) is provided. 

So, there is exist a slightly relationship between A and C

# BC Interaction

According to the plot, the interaction of BC is close to zero which implies the effect of increasing activity(B) does not rely on home visits(C), 

So, B and C are almost independent

# ABC Interaction
According to the plot, the interaction of ABC is very small, and they are almost look like parallel lines.

So, there is not exist a strong relationship between A, B and C.

Hence, in conclusion, to maximize weight loss (kg), the recommended factor settings are:
A = 1: Implement food diary 
B = -1: Decreasing activity
C = 1: Implement home visit 

The   model  

\begin{equation} 
\label{eq1}
\frac{dS}{dt} = -\beta S(t)I(t) 
\end{equation}

The cumulative hazard function H(x), defined by $H(x) = \int_0^x h(u) {\rm d}u$. Thus, for continuous lifetimes, 
\begin{equation}
\label{eq2}
S(x) = \exp[-H(x)] = \exp[-\int_0^x h(u) {\rm d}u]
\end{equation}

\begin{equation}
E[dN_i^*(t)|{Z_i}(t)] = \exp \{ \gamma _0^T{Z_i}(t)\} d{\Lambda _0}(t)
\end{equation}

\begin{align}
\label{eq4}
y = {x^2} + \mathop {\lim }\limits_{n \to \infty } \sum\limits_{i = 1}^n {x_i^2} 
\end{align}





 

 



Markdown is a simple formatting syntax for authoring HTML, PDF, and 
MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.



Here is the table

\begin{table}[htb]
\small
\caption{\scriptsize{very small captions}}
\begin{center}
\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|c|}
\hline
$x$    & 1 & 2   & 3 & 4 & 5    & 6 & 7 & 8    & 9 \\
\hline
$\beta_1$    & 10   & 15  & 25   & 32  &  45 &   51 & 64  & 76  & 98   \\
\hline
\end{tabular}
\end{center}
\label{tab1}
\end{table}

Table \ref{tab1} reports

Recall that
\begin{align}
\label{eq1}
(x+y)^n &=\sum\limits_{k=0}^n \binom{n}{k} x^k y^{n-k}  \\
        &= x^n + n x y^{n-1} +...+nx^{n-1} y+ y^n \nonumber
\end{align}
Equation \eqref{eq1} is called the binomial (Newton) equation. 


The results are
\begin{itemize}

\item A

\item B

\item C

\end{itemize}






in Equation \eqref{eq4}, we see that 

 

 
 
\textbf{References}



# Appendix {-}

The following R code was used for the analysis.

```{r appendix_code, echo=TRUE, eval=FALSE}
######################### Part 1 ###########################
# Load package and data
library(scidesignR)
data(fertdat)

# Overview of dataset
summary(fertdat)

# Separate groups
group_A <- fertdat$fert[fertdat$trt == "A"]
group_B <- fertdat$fert[fertdat$trt == "B"]

# Summary statistics
summary(group_A)
summary(group_B)

# Boxplot
boxplot(fert ~ trt, data = fertdat,
        xlab = "Fertilizer Type",
        ylab = "Wheat Yield",
        main = "Comparison of Wheat Yield by Fertilizer Type")

# Fit model
model <- lm(fert ~ trt, data = fertdat)

# Residuals
res <- residuals(model)

# Normality check
qqnorm(res)
qqline(res)
shapiro.test(res)

# Equal variance check
plot(fitted(model), res,
     xlab = "Fitted Values",
     ylab = "Residuals",
     main = "Residuals vs Fitted Plot")
abline(h = 0, col = "red")
var.test(group_A, group_B)

# Two-sample t-test
t.test(group_A, group_B,
       var.equal = TRUE,
       alternative = "two.sided")

# Key values
mean(group_A)
mean(group_B)
mean(group_A) - mean(group_B)

#################### Part 2 #################################

### Effect Size
library(scidesignR)
data("fertdat")

fert_A <- fertdat$fert[fertdat$trt == "A"]
fert_B <- fertdat$fert[fertdat$trt == "B"]

mean_A <- mean(fert_A)
mean_B <- mean(fert_B)
delta <- mean_A - mean_B

sd_A <- sd(fert_A)
sd_B <- sd(fert_B)

sp <- sqrt(((length(fert_A) - 1) * sd_A^2 + (length(fert_B) - 1) * sd_B^2) / 
             (length(fert_A) + length(fert_B) - 2))

ES <- delta / sp

### Exact Method
power <- 0.8
alpha <- 0.05

exact_n <- power.t.test(
  power = power,
  delta = delta,
  sd = sp, 
  sig.level = alpha
)
exact_n
# Round up the sample size
req_n <- ceiling(exact_n$n)
req_n

### Simulation Power check 
set.seed(1)
alpha <- 0.05

pvals <- replicate(10000,
                   t.test(
                     rnorm(req_n, mean = mean_A, sd = sd_A),
                     rnorm(req_n, mean = mean_B, sd = sd_B),
                     var.equal = TRUE)$p.value)
sim_p <- sum((pvals <= alpha) / 10000)
sim_p

### When the effect side is smaller
delta_small <- 1
power <- 0.8
alpha <- 0.05

exact_n_smallD <- power.t.test(
  power = power,
  delta = delta_small,
  sd = sp, 
  sig.level = alpha
)
exact_n_smallD
# Round up the sample size
req_n_smallD <- ceiling(exact_n_smallD$n)
req_n_smallD

### When the variability is larger
SD_Large <- 10
power <- 0.8
alpha <- 0.05

exact_n_largeSD <- power.t.test(
  power = power,
  delta = delta_small,
  sd = SD_Large, 
  sig.level = alpha
)
exact_n_largeSD
# Round up the sample size
req_n_largeSD <- ceiling(exact_n_largeSD$n)
req_n_largeSD

####################### Part 3 ###############################

```{r, echo=TRUE, eval=FALSE}

# 3A
data(painstudy)
aov.pain <- aov(pain ~ trt, data = painstudy)
summary(aov.pain)

plot(aov.pain$fitted.values, aov.pain$residuals)
qqnorm(aov.pain$residuals)

TukeyHSD(aov.pain)

# 3B
data(shoedat_obs)

shoe_long <- data.frame(
  boy = rep(shoedat_obs$boy, 2),
  treatment = rep(c("A", "B"), each = nrow(shoedat_obs)),
  side = c(shoedat_obs$sideA, shoedat_obs$sideB),
  wear = c(shoedat_obs$wearA, shoedat_obs$wearB)
)

shoe_long$boy <- factor(shoe_long$boy)
shoe_long$treatment <- factor(shoe_long$treatment)

rbd_model <- aov(wear ~ treatment + boy, data = shoe_long)
summary(rbd_model)

unblocked_model <- aov(wear ~ treatment, data = shoe_long)
summary(unblocked_model)

# 3C
data(BR_LatSq)

BR_LatSq$Dogs <- as.factor(BR_LatSq$Dogs)
BR_LatSq$Weeks <- as.factor(BR_LatSq$Weeks)
BR_LatSq$Treat <- as.factor(BR_LatSq$Treat)

lat.model <- aov(Value ~ Treat + Dogs + Weeks, data = BR_LatSq)
summary(lat.model)

```

##################### part 4 ################################
## Load data
wtlossdat <- data.frame(
  A = c(-1, -1, -1, -1,  1,  1,  1,  1),
  B = c(-1, -1,  1,  1, -1, -1,  1,  1),
  C = c(-1,  1, -1,  1, -1,  1, -1,  1),
  y = c(1.1, 1.8, -0.3, -1.1, 1.0, 2.6, -0.4, 0.4)
)

# package and data
library(FrF2)
data(wtlossdat)

# Overview of data
summary(wtlossdat)

# Variables explanation
 A = Food diary (1, -1)
 B = Increasing activity (1, -1)
 C = Home visit (1, -1)
 y = Weight loss (kg)
 
# Method 1: Effects from linear model
Formula used: y ~ A*B*C
Effect = 2 * coefficient (for -1/+1 values)

# Method 2: Effects from mean differences
high = for 1 value
low = for -1 value
Formula used: Effect = mean(high) - mean(low)

# Main effects plots
xlab = x-axis label (factor level)
ylab = y-axis label (mean weight loss)
type = "b" = both points and lines
lwd: line width 
pch: point shape 
col: color

# Two-way interaction plots 
AB: there is not exist a strong relationship between A and B
AC: there is exist a slightly relationship between A and C
BC: B and C are almost independent 

# Three-way interaction plots 
ABC for c = 1/-1 : there is not exist a strong relationship between A, B and C.

